---
title: "individual2"
author: "Brandon Xicon"
date: "2/5/2020"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Tips at Restaurants

In this homework we will use the `tips` data set. This data set is part of the `reshape2` package. You can load the data set by executing the command:
```{r}
#install.packages('reshape2')
data(tips, package="reshape2")
```
If you do not have available the package `reshape2`, issue `install.packages('reshape2')` to install it. The information contained in the data is collected by one waiter, who recorded over the course of a season information about each tip he received working in one restaurant. See `?tips` for a description of all of the variables. 

**Submission instructions**: Create a folder named `ds202_hw3`, and name the RMarkDown file `hw3.Rmd`  which should include your solutions under this folder. For submission, create a GitHub repository named `ds202_hw3` under your GitHub account, and push both `hw3.Rmd` and the knitted `hw3.html` before the deadline. I will assume you use the same GitHub username as for your HW2 submission. The latest version of your homework appearing on GitHub before the deadline is graded. *It is crucial to follow the exact folder structure and file names*, so that your homework can be reproduced and graded by an automated script. This homework is **due on Feb 19** before class.

1.  How many parties did the waiter serve? Store the number in `numParty` and print. 
```{r}
numParty<-nrow(tips)
numParty
```
1. What are the types of variables in the data frame `tips`? Include the code and also explain verbally.
```{r}
sapply(tips, typeof)

```
The types are above, This shows that everything is either a double or an int which includes those that seem to be a String as those are stored as an int in the data and then converted to a string for the client. This however is sort of wrong as the categorical variables are stored as factors. 
1. Create a vector named `day` containing the day of visits in `tips`. The factor levels should be ordered from Thursday to Sunday. Print the variable.
```{r}
day<-factor(tips$day)
day=factor(day,levels(day)[c(4,1,2,3)])
day
```
1. Create a data frame named `female5` containing the meal paid by a female payer in a party with size greater than or euqal to 5. Print the data frame.
```{r}
isfilled<-c(FALSE)
for(i in 1:244){
  if(tips$size[i]>4){
    if(tips$sex[i]=="Female"){
      if(!isfilled){
        female5<-data.frame(tips[i,])
        isfilled<-c(TRUE)
      }
      female5<-rbind(female5,tips[i,])
    }
  }
}

female5

```
1. How does the tipping amount (`tip`) depend on the overall bill (`total_bill`)? Use the `ggplot2` package to make a chart. Describe the relationship in words.
```{r}
library(ggplot2)
ggplot(tips,aes(x=tips$total_bill, y=tips$tip))+geom_point()+geom_smooth(method='lm')+xlab('total Bill ($)')+ylab('tip ($)')
```
1. Describe at least two types of anomalies in the previous plot. What do they mean?
There are outliers such as the over $5 tip on a less than $10 total bill or less $1.25 on an over $30 meal meaning that people can be more generous or evil with tips. 

1. Introduce a variable `tiprate` into the data set which stands for the rate of tips. What is the average rate for tips? Explain verbally. 
```{r}
for(i in 1:numParty){
  tiprate<-c((tips$tip/(tips$total_bill-tips$tip))*100)
}
tips<-cbind(tips,tiprate)
summary(tips)
```
The average rate for tips is around 20% of the bill which is around the preffered bill rate proposed today being around 15-20%.
1. Make a boxplot of the tip rate by time. The x-axis should be ordered by lunch and then dinner. Use `ggplot2` to make a chart. Verbally explain the chart.
```{r}

ggplot(tips,aes(x=reorder(tips$time, +tiprate), y=tips$tiprate))+geom_boxplot()+xlab('time')+ylab('tiprate (%)')

```
Dinner and lunch are around the same as a box plot however has way more outliers shown including the one of 250%

1. How does smoking behavior and gender of the person who pays impact the relationship between tip and total bill? Find a visualization that incorporates all four variables. Interpret the result.
```{r}
ggplot(tips,aes(x=tips$total_bill,y=tips$tip, colour=tips$sex, shape=tips$smoker))+geom_point()+labs(
  x="total bill",
  y="tip",
  colour="sex",
  shape="smoker"
)
```
Men seem to pay more whent he bill is higher the regression seems positively linear nomatter who the person is.

1. Use ggplot2 to find a graphical summary of the relationship between day of the week and gender of the person paying the bill. What can you say about this relationship?
```{r}
ggplot(tips,aes(x=tips$day, fill=tips$sex))+geom_bar(position = "dodge")+labs(
  x="Day",
  y="Count",
  fill="sex"
)
```
Men seem to pay more frequently on saturday and sunday however friday and thursday are around even in pay frequency. 
Note: your submission is supposed to be fully reproducible, i.e. the TA and I will 'knit' your submission in RStudio. 
